%  Copyright (C) 2002 Regents of the University of Michigan, portions used with permission 
%  For more information, see http://csem.engin.umich.edu/tools/swmf
\renewcommand{\bibname}{References}
\begin{thebibliography}{33}

\addcontentsline{toc}{chapter}{\protect References}

\section*{Papers on Numerics and Methods}
\addcontentsline{toc}{section}{\protect Papers on Numerics and Methods}

\bibitem{balsara-spicer}
    D. S. Balsara and D. S. Spicer,
    A Staggered Mesh Algorithm Using High Order Godunov Fluxes to Ensure
	   Solenoidal Magnetic Fields in Magnetohydrodynamic Simulations,
    {\it J. Comput. Phys.} {\bf 149}, 270 (1999).
\bibitem{brackbill-barnes}
    J.\ U.\ Brackbill and D.\ C.\ Barnes,
    The effect of nonzero $\divB$ on the numerical solution 
    of the magnetohydrodynamic equations,
    {\it J.\ Comput.\ Phys.} {\bf 35}, 426 (1980).
\bibitem{dai-woodward}
    W. Dai and P. R. Woodward,
    A simple finite difference scheme for multidimensional
	   magnetohydrodynamic equations,
    {\it J. Comput. Phys.} {\bf 142}, 331 (1998).
\bibitem{evans-hawley} C. R. Evans and J. F. Hawley,
   Simulation of magnetohydrodynamic flows:
	  a constrained transport method,
   {\it Astrophys. J.} {\bf 332}, 659 (1988).
\bibitem{fletcher}
    C.\ A.\ J.\ Fletcher, 
    {\it Computational Techniques for Fluid Dynamics}
    (Springer, Berlin, 1991)
\bibitem{gombosi-etal}
    T. I. Gombosi, K. G. Powell and D. L. De Zeeuw, 
    Axisymmetric modeling of cometary mass loading on an adaptively
    refined grid: MHD results,
    {\it J. Geophys. Res.} {\bf 99}, 21,525 (1994).
\bibitem{harten}
    A.\ Harten,
    High resolution schemes for hyperbolic conservation laws,
    {\it J.\ Comput.\ Phys.} {\bf 49}, 357 (1983).
\bibitem{hestenes-stiefel}
    M. R. Hestenes and E. Stiefel,
    Methods of conjugate gradients for solving linear systems,
    {\it J. Res. Natl. Bur. Stand.} {\bf 49}, 409 (1954).
\bibitem{hirsch}
    C.\ Hirsch,
    {\it Numerical Computation of Internal and External Flows}
    (Wiley, New York, 1990)
\bibitem{lax-wendroff}
    P. D. Lax and B. Wendroff,
    Systems of conservation laws,
    {\it Comm. Pure Appl. Math.} {\bf 13}, 217 (1960).
\bibitem{leveque}
    R.\ J.\ LeVeque,
    {\it Numerical Methods for Conservation Laws}
    (Birkh\"auser, Basel, 1992)
\bibitem{powell}
    K. G. Powell, 
    An approximate Riemann solver for magnetohydrodynamics
    (that works in more than one dimension),
    {\it ICASE Report No 94-24, Langley, VA} (1994).
\bibitem{roe}
    P.\ L.\ Roe,
    Approximate Riemann solvers, parameter vectors, 
    and difference schemes,
    {\it J.\ Comput.\ Phys.} {\bf 43}, 357 (1981).
\bibitem{roe-balsara}
    P. L. Roe and D. S. Balsara, 
    Notes on the eigensystem of magnetohydrodynamics,
    {\it SIAM J. Appl. Math.} {\bf 56}, 57 (1996).\
\bibitem{stone}
    J.\ M.\ Stone and M.\ L.\ Norman, 
    ZEUS-2D: A radiation magnetohydrodynamics code for 
    astrophysical flows in two space dimensions.
    II. The magnetohydrodynamic algorithms and tests,
    {\it Astrophys.\ J. Suppl.} {\bf 80}, 791 (1992).
\bibitem{toth-divb}
    G. T\'oth,
    The $\divB=0$ Constraint in Shock-Capturing Magnetohydrodynamics Codes,
    {\it J. Comput. Phys.} {\bf 161}, 605 (2000).
\bibitem{toth-vac}
    G. T\'oth,
    A general code for modeling MHD flows on parallel computers:
    Versatile Advection Code,
    {\it Astrophys. Lett. \& Comm.} {\bf 34}, 245 (1996).
\bibitem{toth-odstrcil}
    G. T\'oth and D. Odstr\v cil,
    Comparison of some Flux Corrected Transport
    and Total Variation Diminishing Numerical Schemes
    for Hydrodynamic and Magnetohydrodynamic Problems,
    {\it J. Comput. Phys.} {\bf 128}, 82 (1996).
\bibitem{vandervorst}
    H. A. van der Vorst,
    Bi-CGSTAB: A fast and smoothly converging variant of Bi-CG
    for the solution of nonsymmetric linear systems,
    {\it SIAM J. Sci. Statist. Comput.} {\bf 13}, 631 (1992).
\bibitem{vanleer}
    B.\ van Leer,
    Towards the ultimate conservative difference scheme. 
    V. A Second order sequel to Godunov's method,
    {\it J.\ Comput.\ Phys.} {\bf 32}, 101 (1979).
\bibitem{vinokur}
    M. Vinokur,
    A rigorous derivation of the MHD equations based only on 
    Faraday's and Amp\`ere's law,
    presentation at LANL MHD Workshop (1996).
\bibitem{yee}
    H.\ C.\ Yee,
    A class of high-resolution explicit and implicit shock-capturing methods,
    {\it NASA TM-101088} (1989).

\end{thebibliography}


